Individual heterogeneity in the functional network topography of the medial prefrontal cortex

### 2019-04-24 Manuscript.Rmd
halfway through the methods

For all analyses, we defined our search space based on the 17-network parcellation proposed by Yeo et al. [-@Yeo2011]. First, we selected vertices on the cortical surface that were contained by the DN and limbic networks along the medial surface in HCP's 32,000 surface space (fs_LR_32k). Next, we overlaid those networks on a parcellated atlas of the human cortical surface (360 regions) [@Glasser2016], and retained a set of parcels that covered approximately the same brain regions. This resulted in a search space that consisted of 40 parcels across hemispheres (Supplementary Table 1). The search space was naturally divided into two spatially discontinuous clusters that covered PCC and mPFC, facilitating the study of each general region separately.

@@

We used a novel approach to cortical surface parcel-based meta-analysis to assess whether individual parcels within the search space were preferentially associated with valuation or DN effects. For subjective valuation, we gathered peak brain coordinates of activity from 200 studies that were associated with positive effects for current or prospective rewards [@Bartra2013]. For DN, we acquired coordinates from a previous meta-analysis of 80 studies that were related to task-deactivation [@Laird2009]. These coordinates represent the areas that exceeded the statistical significance threshold in each original study. For each study, we created an indicator map in standard space (MNI152, 1 mm resolution) which contained values of 1 in a 10 mm radius sphere around each reported activation peak, and values of 0 elsewhere [@Wager2009]. The indicator map for each study was then projected to a standard cortical mesh (fsaverage, 160,000 vertices, projfrac-max from 0 to 1 by 0.25) using FreeSurfer's mri_vol2surf [@Fischl1999; @Dale1999] (http://surfer.nmr.mgh.harvard.edu/). **TODO: The Glasser et al. (2016) parcellation was resampled to fsaverage to pair spheres to parcels from our search space**. Any sphere sections that overlapped with the cortex (no lower limit restrictions) were mapped to the corresponding parcel, and we tallied how many studies had positive indicator values in each parcel. Two studies from the subjective valuation corpus were removed because they did not contain cortical voxels, leaving a final number of 198 studies.

With these lists, we tested for parcels that were significantly more strongly associated with one domain than the other with a chi-squared test. We permuted the label of every study (i.e. DN or valuation) 5000 times while preserving the total number of studies observed in each domain, and on each iteration stored the maximum resulting statistic (across all parcels) from per-parel chi-squared tests comparing the proportion of studies with activation in that parcel in the two domains. This gave us a null distribution of 5000 maximum chi-squared values. The 95th percentile of this distribution served as an FWE-corrected significance threshold to evaluate unpermuted chi-squared values.

@@

Data initially underwent the HCP minimal preprocessing pipeline [@Glasser2013], which included gradient nonlinearity correction, motion correction, EPI distortion correction, high-pass filtering (0.0005 Hz threshold), MNI152-based normalization, surface reconstruction, and mapping of functional data to the surface (details can be found on Glasset et al., 2013). In addition, data underwent temporal denoising based on independent components [FIX; @Griffanti2014; @Salimi-Khorshidi2014]. Data were further preprocessed using an in-house pipeline described previously [@Tobyne2017]. Steps included linear interpolation across high motion timepoints with over 0.5 mm of framewise displacement, band-pass filtering (allowed frequencies ranged from 0.009 and 0.08 Hz), temporal denoising via grayordinate signal regression [@Burgess2016] and censoring of high-motion timepoints by deletion. Finally, data acquired on the same day (i.e. left-right and right-left phase encoding session) were temporally demeaned. The post-processed time series had a median of `r median(corrCounts$nTR)` time points (minimum = `r min(corrCounts$nTR)`) across participants. Each subject's brain was comprised of 32k standard grayordinates per hemisphere (combined in a CIFTI file). We retained only the cortical surfaces, which resulted in 59,412 total surface vertices per subject.

@@

All network analyses were performed using the igraph package [@G.2006; https://igraph.org/r/] in R [v. 3.4.1; @RTeam2018; https://www.r-project.org/]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, in which the nodes were surface vertices and the edges were the correlations among them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify significant connections. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were set to 0 (all other edges retained their respective z-values). The proportion of remaining edges (i.e. network density) was high for all individuals (mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`), which were mostly composed of positive correlations (positive: mean = `r mPos`, sd = `r sdPos`; negative: mean = `r mNeg`, sd = `r sdNeg`). Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Values of 1 of this transformed distribution were reset to 0 in order to avoid the inclusion of these edges in the network analyses. We generated and analyzed network weight matrices at four levels: (1) for each run separately (1200 TRs each); (2) for the concatenated time series for the two runs on each day (2400 TRs total); (3) for each subject's full dataset (4800 TRs total); and (4) on each step of a sliding window analysis (see Partition Evaluation for more details).

>>>>

For all analyses, we defined our search space based on the 17-network parcellation proposed by Yeo et al. [-@Yeo2011]. First, we selected vertices on the medial cortical surface that were contained by the DN and limbic networks in HCP's 32,000 surface space (fs_LR_32k). Next, we overlaid those networks on a parcellated atlas of the human cortical surface (360 regions) [@Glasser2016], and retained a set of parcels that covered approximately the same brain regions. This resulted in a search space that consisted of 40 parcels across hemispheres (Supplementary Table 1). The search space was naturally divided into two spatially non-contiguous clusters that covered PCC and mPFC, facilitating the examination of each general region separately.

@@

We used a novel approach to cortical surface parcel-based meta-analysis to assess whether individual parcels within the search space were preferentially associated with valuation or DN effects. For subjective valuation, we gathered peak brain coordinates of activity from 200 studies that were associated with positive effects in contrasts of high-versus-low-value outcomes or prospects [@Bartra2013]. For DN, we acquired coordinates from 80 studies that were related to task-deactivation [@Laird2009]. These coordinates represent the areas that exceeded the statistical significance threshold in each original study. For each study, we created an indicator map in standard space (MNI152, 1 mm resolution) which contained values of 1 in a 10 mm radius sphere around each reported activation peak, and values of 0 elsewhere [@Wager2009]. The indicator map for each study was then projected to a standard cortical mesh (fsaverage, 160,000 vertices, projfrac-max from 0 to 1 by 0.25) using FreeSurfer's mri_vol2surf [@Fischl1999; @Dale1999] (http://surfer.nmr.mgh.harvard.edu/). We then resampled the Glasser et al. (2016) parcellation to fsaverage, and tallied how many studies had positive indicator values intersecting with each cortical parcel (the details of the resampling procedure are described in https://wiki.humanconnectome.org/display/PublicData/HCP+Users+FAQ#HCPUsersFAQ-9.HowdoImapdatabetweenFreeSurferandHCP, and were implemented using a custom script https://github.com/stobyne/Spherical-Surface-Swapper). Two studies from the subjective valuation corpus were removed because they did not contain cortical voxels, leaving a final number of 198 studies.

To test for parcels that were significantly more strongly associated with one domain than the other, we performed per-parcel chi-squared tests comparing the proportion of studies with activation in that parcel between the two domains. We permuted the study domain labels (i.e. DN or valuation) 5000 times while preserving the total number of studies in each domain, and on each iteration stored the maximum resulting chi-squared statistic (across all parcels). This gave us a null distribution of 5000 maximum chi-squared values. The 95th percentile of this distribution served as an FWE-corrected significance threshold to evaluate unpermuted chi-squared values.

@@

Data initially underwent the HCP minimal preprocessing pipeline [@Glasser2013], which included gradient nonlinearity correction, motion correction, EPI distortion correction, high-pass filtering (0.0005 Hz threshold), MNI152-based normalization, surface reconstruction, and mapping of functional data to the a standardized cortical surface model (details can be found on Glasser et al., 2013). In addition, data underwent temporal denoising based on independent components [FIX; @Griffanti2014; @Salimi-Khorshidi2014]. Data were further preprocessed using an in-house pipeline described previously [@Tobyne2017]. Steps included linear interpolation across high motion timepoints with over 0.5 mm of framewise displacement, band-pass filtering (allowed frequencies ranged from 0.009 and 0.08 Hz), temporal denoising via mean grayordinate signal regression [@Burgess2016] and censoring of high-motion timepoints by deletion. Finally, data acquired on the same day (i.e. left-right and right-left phase encoding session) were temporally demeaned. The post-processed time series had a median of `r median(corrCounts$nTR)` time points (minimum = `r min(corrCounts$nTR)`) across participants. Each subject's brain was comprised of 32k standard grayordinates per hemisphere (combined in a CIFTI file). We retained only the cortical surfaces, which resulted in 59,412 total surface vertices per subject.

@@

All network analyses were performed using the igraph package [v. 1.1.2; @G.2006; https://igraph.org/r/] in R [v. 3.4.1; https://www.r-project.org/; @RTeam2018]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, in which the nodes were surface vertices and the edges were the correlations among them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify significant connections. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were set to 0 (all other edges retained their respective z-values). The proportion of remaining edges (i.e. network density) was high for all individuals (mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`), and the retained edges mostly consisted of positive correlations (mean proportion positive = `r mPos`, sd = `r sdPos`). Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution within the set of retained edges. Non-retained edges were reset to 0 to exclude them from the network analyses. We generated and analyzed network weight matrices at four levels: (1) for each subject's full dataset (4800 TRs total); (2) on each step of a sliding window analysis (see Partition Evaluation for more details); (3) for the concatenated time series for the two runs on each day (2400 TRs total); and (4) for each run separately (1200 TRs each).
